---
title: "Notes: Matrices in R"
author: "Ivan Jacob Agaloos Pesigan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Notes: Matrices in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_options, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  error = TRUE,
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

A matrix is a two dimensional data structure in `R`.
The dimensions ($m \times n$)
are stored as an `attribute`
`dim`
which can be accessed using the `attributes()` function.

The following are some useful functions related to matrices.

- `matrix()` - creates a matrix
- `as.matrix()` - attempts to coerce its argument into a matrix
- `is.matrix()` - tests if its argument is a matrix

The example below creates a $3 \times 3$ matrix $\mathbf{A}$
from a vector $\mathbf{a}$
consisting is numbers 1 to 9.

```{r}
a <- 1:9
A <- matrix(
  data = a,
  ncol = 3
)
A
attributes(A)
```

The dimensions are set using `nrow` and `ncol`.
Notice that in the example above
only `ncol` is specified.
`R` is smart enough to determine
the corresponding `nrow`
based on the length of the input data.

By default,
`matrix()`
is populated by column,
setting the argument `byrow = TRUE`
populates the matrix by row.

```{r}
b <- 1:9
B <- matrix(
  data = b,
  ncol = 3,
  byrow = TRUE
)
B
```

A vector can be transformed into a matrix
by providing the appropriate dimensions
using the `dim()` function.

```{r}
x <- 1:9
x
dim(x) <- c(3, 3)
x
```

Matrices can also be created
by binding several vectors
by column (`cbind()`)
or
by row (`rbind()`).

```{r}
col_1 <- 1:9
col_2 <- 10:18
col_3 <- 19:27
C <- cbind(col_1, col_2, col_3)
C
is.matrix(C)
row_1 <- 1:9
row_2 <- 10:18
row_3 <- 19:27
R <- rbind(row_1, row_2, row_3)
R
is.matrix(R)
```

Notice that dimension names are inherited
from the vectors used to create the matrix.
Dimension names can also be explicitly provided
using the argument `dimnames`
which accepts a list.
The first element of the list provides row names.
The second element of the list provides column names.

```{r}
X <- matrix(
  data = 1:9,
  ncol = 3,
  dimnames = list(
    c("a", "b", "c"),
    c("x", "y", "z")
  )
)
X
```

The functions `colnames()` and `rownames()`
can also be used to name dimensions of an existing matrix.

```{r}
colnames(X) <- c("col_1", "col_2", "col_3")
rownames(X) <- c("row_1", "row_2", "row_3")
X
```

A data frame can be coerced into a matrix using `as.matrix()`.
Note that unlike data frames,
matrices in `R` can only contain values of the same type.
Elements are coerced to a common type.

```{r}
Y <- as.matrix(iris)
head(Y)
```

Since the data frame iris has a column with character strings (column 5),
all the elements in the resulting matrix `Y`
are coerced to the type `character`.

Since we are only interested in numeric matrices,
removing the fifth column and coercing the data frame
into a matrix results in a numeric matrix
consisting of doubles.

```{r}
df <- iris[-c(5)]
head(df)
is.matrix(df)
is.data.frame(df)
Z <- as.matrix(df)
head(Z)
is.matrix(Z)
typeof(Z)
```

## Matrix Operations

```{r, echo = FALSE}
m <- 3
n <- 3
x <- 1:9
a <- sample(
  x = x,
  size = m * n,
  replace = TRUE
)
b <- sample(
  x = x,
  size = m * n,
  replace = TRUE
)
c <- sample(
  x = x,
  size = m * n,
  replace = TRUE
)
k <- sample(
  x = x,
  size = 1
)
s <- sample(
  x = x,
  size = 1
)
A <- matrix(
  data = a,
  ncol = m
)
B <- matrix(
  data = b,
  ncol = m
)
C <- matrix(
  data = c,
  ncol = m
)
```

Let

\begin{equation}
  k
  =
  `r k` ,
\end{equation}

\begin{equation}
  s
  =
  `r s`,
\end{equation}

\begin{equation}
  \mathbf{A}
  =
  \begin{bmatrix}
    `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
    `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
    `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]`
  \end{bmatrix} ,
\end{equation}

\begin{equation}
  \mathbf{B}
  =
  \begin{bmatrix}
    `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
    `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
    `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]`
  \end{bmatrix} ,
\end{equation}

\noindent and

\begin{equation}
  \mathbf{C}
  =
  \begin{bmatrix}
    `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
    `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
    `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]`
  \end{bmatrix} .
\end{equation}

### Scalar multiplication

The operator `*` is used for scalar multiplication.

\begin{equation}
  k \mathbf{A}
\end{equation}

```{r}
k * A
```

\begin{equation}
  s \left( k \mathbf{A} \right)
  =
  \left( sk \mathbf{A} \right)
  =
  \left( ks \mathbf{A} \right)
  =
  k \left( s \mathbf{A} \right) .
\end{equation}

```{r}
s * (k * A)
(s * k * A)
(k * s) * A
k * (s * A)
```

### Matrix addition and subtraction

The operator `+` is used for matrix addition.
The operator `-` is used for matrix subtraction.

\begin{equation}
  \mathbf{A} + \mathbf{B}
\end{equation}

```{r}
A + B
A - B
```

\begin{equation}
  \mathbf{A} + \mathbf{B}
  =
  \mathbf{B} + \mathbf{A}
\end{equation}

```{r}
A + B
B + A
```

\begin{equation}
  \mathbf{A}
  +
  \left(
    \mathbf{B}
    +
    \mathbf{C}
  \right)
  =
  \left(
    \mathbf{A}
    +
    \mathbf{B}
  \right)
  +
  \mathbf{C}
  =
  \mathbf{A}
  +
  \mathbf{B}
  +
  \mathbf{C}
\end{equation}

```{r}
A + (B + C)
(A + B) + C
A + B + C
```

\begin{equation}
  s
  \left(
    \mathbf{A}
    +
    \mathbf{B}
  \right)
  =
  s \mathbf{A}
  +
  s \mathbf{B}
\end{equation}

```{r}
s * (A + B)
s * A + s * B
```

\begin{equation}
  \left(
    s
    +
    k
  \right)
  \mathbf{A}
  =
  s
  \mathbf{A}
  +
  k
  \mathbf{A}
\end{equation}

```{r}
(s + k) * A
s * A + k * A
```




```{r, echo = FALSE}
# do not change m and n
m <- 4
n <- 3
a <- c(
  6,
  5,
  9,
  -2,
  10,
  1,
  7,
  11,
  4.25,
  12,
  -8,
  3
)
b <- c(
  3,
  10,
  -4,
  6,
  5.5,
  9,
  -1,
  7,
  12,
  2,
  8.01,
  11
)
c <- c(
  9,
  5,
  3,
  1.78,
  10,
  7,
  8,
  12,
  2.45,
  6,
  11,
  4
)
k <- 2
s <- 3
A <- matrix(
  data = a,
  nrow = m,
  ncol = n
)
B <- matrix(
  data = b,
  nrow = m,
  ncol = n
)
C <- matrix(
  data = c,
  nrow = m,
  ncol = n
)
```



```{r}
k * A
```


```{r}
A + B
```



## Examples

Let
\begin{equation}
  s
  =
  `r s` ,
\end{equation}

\begin{equation}
  k
  =
  `r k` ,
\end{equation}

\noindent and

\begin{equation}
  \mathbf{A}
  =
  \begin{bmatrix}
    `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
    `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
    `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
    `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
  \end{bmatrix}
\end{equation}

1. $k \mathbf{A}$

    \begin{equation}
    \begin{split}
      k \mathbf{A}
      &=
      `r k`
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r k * A[1, 1]` & `r k * A[1, 2]` &  `r k * A[1, 3]` \\
        `r k * A[2, 1]` & `r k * A[2, 2]` &  `r k * A[2, 3]` \\
        `r k * A[3, 1]` & `r k * A[3, 2]` &  `r k * A[3, 3]` \\
        `r k * A[4, 1]` & `r k * A[4, 2]` &  `r k * A[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
k * A
    ```

1. $\mathbf{A} + \mathbf{B}$

    \begin{equation}
    \begin{split}
      \mathbf{A} + \mathbf{B}
      &=
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
        `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
        `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
        `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1]` & `r A[1, 2] + B[1, 2]` & `r A[1, 3] + B[1, 3]` \\
        `r A[2, 1] + B[2, 1]` & `r A[2, 2] + B[2, 2]` & `r A[2, 3] + B[2, 3]` \\
        `r A[3, 1] + B[3, 1]` & `r A[3, 2] + B[3, 2]` & `r A[3, 3] + B[3, 3]` \\
        `r A[4, 1] + B[4, 1]` & `r A[4, 2] + B[4, 2]` & `r A[4, 3] + B[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
A + B
    ```
1. $s \left( k \mathbf{A} \right)$

    \begin{equation}
    \begin{split}
      s
      \left(
        k \mathbf{A}
      \right)
      &=
      `r s`
      \left(
        `r k`
        \begin{bmatrix}
          `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
          `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
          `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
          `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
        \end{bmatrix}
      \right) \\
      &=
      `r s`
      \begin{bmatrix}
        `r k * A[1, 1]` & `r k * A[1, 2]` & `r k * A[1, 3]` \\
        `r k * A[2, 1]` & `r k * A[2, 2]` & `r k * A[2, 3]` \\
        `r k * A[3, 1]` & `r k * A[3, 2]` & `r k * A[3, 3]` \\
        `r k * A[4, 1]` & `r k * A[4, 2]` & `r k * A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r s * k * A[1, 1]` & `r s * k * A[1, 2]` & `r s * k * A[1, 3]` \\
        `r s * k * A[2, 1]` & `r s * k * A[2, 2]` & `r s * k * A[2, 3]` \\
        `r s * k * A[3, 1]` & `r s * k * A[3, 2]` & `r s * k * A[3, 3]` \\
        `r s * k * A[4, 1]` & `r s * k * A[4, 2]` & `r s * k * A[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
s * (k * A)
    ```

1. $\left( sk \mathbf{A} \right)$

    \begin{equation}
    \begin{split}
      \left(
        sk
        \mathbf{A}
      \right)
      &=
      \left(
        `r s` \times `r k`
        \begin{bmatrix}
          `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
          `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
          `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]`
        \end{bmatrix}
      \right) \\
      &=
      `r s * k`
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r s * k * A[1, 1]` & `r s * k * A[1, 2]` & `r s * k * A[1, 3]` \\
        `r s * k * A[2, 1]` & `r s * k * A[2, 2]` & `r s * k * A[2, 3]` \\
        `r s * k * A[3, 1]` & `r s * k * A[3, 2]` & `r s * k * A[3, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
(s * k * A)
    ```

1. $\left( ks \mathbf{A} \right)$

    \begin{equation}
    \begin{split}
      \left(
        ks
        \mathbf{A}
      \right)
      &=
      \left(
        `r k` \times `r s`
        \begin{bmatrix}
          `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
          `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
          `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
          `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
        \end{bmatrix}
      \right) \\
      &=
      `r k * s`
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r k * s * A[1, 1]` & `r k * s * A[1, 2]` & `r k * s * A[1, 3]` \\
        `r k * s * A[2, 1]` & `r k * s * A[2, 2]` & `r k * s * A[2, 3]` \\
        `r k * s * A[3, 1]` & `r k * s * A[3, 2]` & `r k * s * A[3, 3]` \\
        `r k * s * A[4, 1]` & `r k * s * A[4, 2]` & `r k * s * A[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
(k * s * A)
    ```

1. $k \left( s \mathbf{A} \right)$

    \begin{equation}
    \begin{split}
      k
      \left(
        s
        \mathbf{A}
      \right)
      &=
      `r k`
      \left(
        `r s`
        \begin{bmatrix}
          `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
          `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
          `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
          `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
        \end{bmatrix}
      \right) \\
      &=
      `r k`
      \begin{bmatrix}
        `r s * A[1, 1]` & `r s * A[1, 2]` & `r s * A[1, 3]` \\
        `r s * A[2, 1]` & `r s * A[2, 2]` & `r s * A[2, 3]` \\
        `r s * A[3, 1]` & `r s * A[3, 2]` & `r s * A[3, 3]` \\
        `r s * A[4, 1]` & `r s * A[4, 2]` & `r s * A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r k * s * A[1, 1]` & `r k * s * A[1, 2]` & `r k * s * A[1, 3]` \\
        `r k * s * A[2, 1]` & `r k * s * A[2, 2]` & `r k * s * A[2, 3]` \\
        `r k * s * A[3, 1]` & `r k * s * A[3, 2]` & `r k * s * A[3, 3]` \\
        `r k * s * A[4, 1]` & `r k * s * A[4, 2]` & `r k * s * A[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
k * (s * A)
    ```

1. $\left( -1 \right) \mathbf{A} = - \mathbf{A}$

    \begin{equation}
    \begin{split}
      \left( -1 \right) \mathbf{A}
      &=
      -1
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r -1 * A[1, 1]` & `r -1 * A[1, 2]` & `r -1 * A[1, 3]` \\
        `r -1 * A[2, 1]` & `r -1 * A[2, 2]` & `r -1 * A[2, 3]` \\
        `r -1 * A[3, 1]` & `r -1 * A[3, 2]` & `r -1 * A[3, 3]` \\
        `r -1 * A[4, 1]` & `r -1 * A[4, 2]` & `r -1 * A[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
(-1) * A
    ```

1. $\mathbf{A} + \mathbf{B} = \mathbf{B} + \mathbf{A}$

    \begin{equation}
    \begin{split}
      \mathbf{A} + \mathbf{B}
      &=
      \mathbf{B} + \mathbf{A} \\
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
        `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
        `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
        `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
      \end{bmatrix}
      &=
      \begin{bmatrix}
        `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
        `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
        `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
        `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1]` & `r A[1, 2] + B[1, 2]` & `r A[1, 3] + B[1, 3]` \\
        `r A[2, 1] + B[2, 1]` & `r A[2, 2] + B[2, 2]` & `r A[2, 3] + B[2, 3]` \\
        `r A[3, 1] + B[3, 1]` & `r A[3, 2] + B[3, 2]` & `r A[3, 3] + B[3, 3]` \\
        `r A[4, 1] + B[4, 1]` & `r A[4, 2] + B[4, 2]` & `r A[4, 3] + B[4, 3]`
      \end{bmatrix}
      &=
      \begin{bmatrix}
        `r B[1, 1] + A[1, 1]` & `r B[1, 2] + A[1, 2]` & `r B[1, 3] + A[1, 3]` \\
        `r B[2, 1] + A[2, 1]` & `r B[2, 2] + A[2, 2]` & `r B[2, 3] + A[2, 3]` \\
        `r B[3, 1] + A[3, 1]` & `r B[3, 2] + A[3, 2]` & `r B[3, 3] + A[3, 3]` \\
        `r B[4, 1] + A[4, 1]` & `r B[3, 2] + A[4, 2]` & `r B[4, 3] + A[4, 3]`
      \end{bmatrix} \\
    \end{split}
    \end{equation}

    ```{r}
A + B
B + A
    ```

1. $\mathbf{A} + \left( \mathbf{B} + \mathbf{C} \right) = \left( \mathbf{A} + \mathbf{B} \right) + \mathbf{C} = \mathbf{A} + \mathbf{B} \mathbf{C}$

    \begin{equation}
      \mathbf{A}
      +
      \left(
        \mathbf{B}
        +
        \mathbf{C}
      \right)
      =
      \left(
        \mathbf{A}
        +
        \mathbf{B}
      \right)
      +
      \mathbf{C}
      =
      \mathbf{A}
      +
      \mathbf{B}
      +
      \mathbf{C}
    \end{equation}

    \begin{equation}
    \begin{split}
      \mathbf{A}
      +
      \left(
        \mathbf{B}
        +
        \mathbf{C}
      \right)
      &=
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      \left(
        \begin{bmatrix}
          `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
          `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
          `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
          `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
        \end{bmatrix}
        +
        \begin{bmatrix}
          `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
          `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
          `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]` \\
          `r C[4, 1]` & `r C[4, 2]` & `r C[4, 3]`
        \end{bmatrix}
      \right) \\
      &=
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r B[1, 1] + C[1, 1]` & `r B[1, 2] + C[1, 2]` & `r B[1, 3] + C[1, 3]` \\
        `r B[2, 1] + C[2, 1]` & `r B[2, 2] + C[2, 2]` & `r B[2, 3] + C[2, 3]` \\
        `r B[3, 1] + C[3, 1]` & `r B[3, 2] + C[3, 2]` & `r B[3, 3] + C[3, 3]` \\
        `r B[4, 1] + C[4, 1]` & `r B[4, 2] + C[4, 2]` & `r B[4, 3] + C[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1] + C[1, 1]` & `r A[1, 2] + B[1, 2] + C[1, 2]` & `r A[1, 3] + B[1, 3] + C[1, 3]` \\
        `r A[2, 1] + B[2, 1] + C[2, 1]` & `r A[2, 2] + B[2, 2] + C[2, 2]` & `r A[2, 3] + B[2, 3] + C[2, 3]` \\
        `r A[3, 1] + B[3, 1] + C[3, 1]` & `r A[3, 2] + B[3, 2] + C[3, 2]` & `r A[3, 3] + B[3, 3] + C[3, 3]` \\
        `r A[4, 1] + B[4, 1] + C[4, 1]` & `r A[4, 2] + B[4, 2] + C[4, 2]` & `r A[4, 3] + B[4, 3] + C[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    \begin{equation}
    \begin{split}
      \left(
        \mathbf{A}
        +
        \mathbf{B}
      \right)
      +
      \mathbf{C}
      &=
      \left(
        \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
        \end{bmatrix}
        +
        \begin{bmatrix}
          `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
          `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
          `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
          `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
        \end{bmatrix}
      \right)
      +
      \begin{bmatrix}
        `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
        `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
        `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]` \\
        `r C[4, 1]` & `r C[4, 2]` & `r C[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1]` & `r A[1, 2] + B[1, 2]` & `r A[1, 3] + B[1, 3]` \\
        `r A[2, 1] + B[2, 1]` & `r A[2, 2] + B[2, 2]` & `r A[2, 3] + B[2, 3]` \\
        `r A[3, 1] + B[3, 1]` & `r A[3, 2] + B[3, 2]` & `r A[3, 3] + B[3, 3]` \\
        `r A[4, 1] + B[4, 1]` & `r A[4, 2] + B[4, 2]` & `r A[4, 3] + B[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
        `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
        `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]` \\
        `r C[4, 1]` & `r C[4, 2]` & `r C[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1] + C[1, 1]` & `r A[1, 2] + B[1, 2] + C[1, 2]` & `r A[1, 3] + B[1, 3] + C[1, 3]` \\
        `r A[2, 1] + B[2, 1] + C[2, 1]` & `r A[2, 2] + B[2, 2] + C[2, 2]` & `r A[2, 3] + B[2, 3] + C[2, 3]` \\
        `r A[3, 1] + B[3, 1] + C[3, 1]` & `r A[3, 2] + B[3, 2] + C[3, 2]` & `r A[3, 3] + B[3, 3] + C[3, 3]` \\
        `r A[4, 1] + B[4, 1] + C[4, 1]` & `r A[4, 2] + B[4, 2] + C[4, 2]` & `r A[4, 3] + B[4, 3] + C[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    \begin{equation}
    \begin{split}
      \mathbf{A}
      +
      \mathbf{B}
      +
      \mathbf{C}
      &=
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
        `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
        `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
        `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
        `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
        `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]` \\
        `r C[4, 1]` & `r C[4, 2]` & `r C[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1]` & `r A[1, 2] + B[1, 2]` & `r A[1, 3] + B[1, 3]` \\
        `r A[2, 1] + B[2, 1]` & `r A[2, 2] + B[2, 2]` & `r A[2, 3] + B[2, 3]` \\
        `r A[3, 1] + B[3, 1]` & `r A[3, 2] + B[3, 2]` & `r A[3, 3] + B[3, 3]` \\
        `r A[4, 1] + B[4, 1]` & `r A[4, 2] + B[4, 2]` & `r A[4, 3] + B[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r C[1, 1]` & `r C[1, 2]` & `r C[1, 3]` \\
        `r C[2, 1]` & `r C[2, 2]` & `r C[2, 3]` \\
        `r C[3, 1]` & `r C[3, 2]` & `r C[3, 3]` \\
        `r C[4, 1]` & `r C[4, 2]` & `r C[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1] + C[1, 1]` & `r A[1, 2] + B[1, 2] + C[1, 2]` & `r A[1, 3] + B[1, 3] + C[1, 3]` \\
        `r A[2, 1] + B[2, 1] + C[2, 1]` & `r A[2, 2] + B[2, 2] + C[2, 2]` & `r A[2, 3] + B[2, 3] + C[2, 3]` \\
        `r A[3, 1] + B[3, 1] + C[3, 1]` & `r A[3, 2] + B[3, 2] + C[3, 2]` & `r A[3, 3] + B[3, 3] + C[3, 3]` \\
        `r A[4, 1] + B[4, 1] + C[4, 1]` & `r A[4, 2] + B[4, 2] + C[4, 2]` & `r A[4, 3] + B[4, 3] + C[4, 3]`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
A + (B + C)
(A + B) + C
A + B + C
    ```

1. $s \left( \mathbf{A} + \mathbf{B} \right) = s \mathbf{A} + s \mathbf{B}$

    \begin{equation}
    \begin{split}
      s
      \left(
        \mathbf{A}
        +
        \mathbf{B}
      \right)
      &=
      `r s`
      \left(
        \begin{bmatrix}
          `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
          `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
          `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
          `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
        \end{bmatrix}
        +
        \begin{bmatrix}
          `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
          `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
          `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
          `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
        \end{bmatrix}
      \right) \\
      &=
      `r s`
      \begin{bmatrix}
        `r A[1, 1] + B[1, 1]` & `r A[1, 2] + B[1, 2]` & `r A[1, 3] + B[1, 3]` \\
        `r A[2, 1] + B[2, 1]` & `r A[2, 2] + B[2, 2]` & `r A[2, 3] + B[2, 3]` \\
        `r A[3, 1] + B[3, 1]` & `r A[3, 2] + B[3, 2]` & `r A[3, 3] + B[3, 3]` \\
        `r A[4, 1] + B[4, 1]` & `r A[4, 2] + B[4, 2]` & `r A[4, 3] + B[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r s * (A[1, 1] + B[1, 1])` & `r s * (A[1, 2] + B[1, 2])` & `r s * (A[1, 3] + B[1, 3])` \\
        `r s * (A[2, 1] + B[2, 1])` & `r s * (A[2, 2] + B[2, 2])` & `r s * (A[2, 3] + B[2, 3])` \\
        `r s * (A[3, 1] + B[3, 1])` & `r s * (A[3, 2] + B[3, 2])` & `r s * (A[3, 3] + B[3, 3])` \\
        `r s * (A[4, 1] + B[4, 1])` & `r s * (A[4, 2] + B[4, 2])` & `r s * (A[4, 3] + B[4, 3])`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
s * (A + B)
    ```

    \begin{equation}
    \begin{split}
      s \mathbf{A}
      +
      s \mathbf{B}
      &=
      `r s`
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      `r s`
      \begin{bmatrix}
        `r B[1, 1]` & `r B[1, 2]` & `r B[1, 3]` \\
        `r B[2, 1]` & `r B[2, 2]` & `r B[2, 3]` \\
        `r B[3, 1]` & `r B[3, 2]` & `r B[3, 3]` \\
        `r B[4, 1]` & `r B[4, 2]` & `r B[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r s * A[1, 1]` & `r s * A[1, 2]` & `r s * A[1, 3]` \\
        `r s * A[2, 1]` & `r s * A[2, 2]` & `r s * A[2, 3]` \\
        `r s * A[3, 1]` & `r s * A[3, 2]` & `r s * A[3, 3]` \\
        `r s * A[4, 1]` & `r s * A[4, 2]` & `r s * A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
        `r s * B[1, 1]` & `r s * B[1, 2]` & `r s * B[1, 3]` \\
        `r s * B[2, 1]` & `r s * B[2, 2]` & `r s * B[2, 3]` \\
        `r s * B[3, 1]` & `r s * B[3, 2]` & `r s * B[3, 3]` \\
        `r s * B[4, 1]` & `r s * B[4, 2]` & `r s * B[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r (s * A[1, 1]) + (s * B[1, 1])` & `r (s * A[1, 2]) + (s * B[1, 2])` & `r (s * A[1, 3]) + (s * B[1, 3])` \\
        `r (s * A[2, 1]) + (s * B[2, 1])` & `r (s * A[2, 2]) + (s * B[2, 2])` & `r (s * A[2, 3]) + (s * B[2, 3])` \\
        `r (s * A[3, 1]) + (s * B[3, 1])` & `r (s * A[3, 2]) + (s * B[3, 2])` & `r (s * A[3, 3]) + (s * B[3, 3])` \\
        `r (s * A[4, 1]) + (s * B[4, 1])` & `r (s * A[4, 2]) + (s * B[4, 2])` & `r (s * A[4, 3]) + (s * B[4, 3])`
      \end{bmatrix}
    \end{split}
    \end{equation}


    ```{r}
s * A + s * B
    ```

1. $\left( s + k \right) \mathbf{A} = s \mathbf{A} + k \mathbf{A}$

    \begin{equation}
    \begin{split}
      \left(
        s
        +
        k
      \right)
      \mathbf{A}
      &=
      \left(
        `r s`
        +
        `r k`
      \right)
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      `r s + k`
      \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
        `r (s + k) * (A[1, 1])` & `r (s + k) * (A[1, 2])` & `r (s + k) * (A[1, 3])` \\
        `r (s + k) * (A[2, 1])` & `r (s + k) * (A[2, 2])` & `r (s + k) * (A[2, 3])` \\
        `r (s + k) * (A[3, 1])` & `r (s + k) * (A[3, 2])` & `r (s + k) * (A[3, 3])` \\
        `r (s + k) * (A[4, 1])` & `r (s + k) * (A[4, 2])` & `r (s + k) * (A[4, 3])`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
(s + k) * A
    ```

    \begin{equation}
    \begin{split}
      s
      \mathbf{A}
      +
      k
      \mathbf{A}
      &=
      `r s`
       \begin{bmatrix}
        `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
        `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
        `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
        `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix}
      +
      `r k`
      \begin{bmatrix}
       `r A[1, 1]` & `r A[1, 2]` & `r A[1, 3]` \\
       `r A[2, 1]` & `r A[2, 2]` & `r A[2, 3]` \\
       `r A[3, 1]` & `r A[3, 2]` & `r A[3, 3]` \\
       `r A[4, 1]` & `r A[4, 2]` & `r A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
       `r s * A[1, 1]` & `r s * A[1, 2]` & `r s * A[1, 3]` \\
       `r s * A[2, 1]` & `r s * A[2, 2]` & `r s * A[2, 3]` \\
       `r s * A[3, 1]` & `r s * A[3, 2]` & `r s * A[3, 3]` \\
       `r s * A[4, 1]` & `r s * A[4, 2]` & `r s * A[4, 3]`
      \end{bmatrix}
      +
      \begin{bmatrix}
       `r k * A[1, 1]` & `r k * A[1, 2]` & `r k * A[1, 3]` \\
       `r k * A[2, 1]` & `r k * A[2, 2]` & `r k * A[2, 3]` \\
       `r k * A[3, 1]` & `r k * A[3, 2]` & `r k * A[3, 3]` \\
       `r k * A[4, 1]` & `r k * A[4, 2]` & `r k * A[4, 3]`
      \end{bmatrix} \\
      &=
      \begin{bmatrix}
       `r (s * A[1, 1]) + (k * A[1, 1])` & `r (s * A[1, 2]) + (k * A[1, 2])` & `r (s * A[1, 3]) + (k * A[1, 3])` \\
       `r (s * A[2, 1]) + (k * A[2, 1])` & `r (s * A[2, 2]) + (k * A[2, 2])` & `r (s * A[2, 3]) + (k * A[2, 3])` \\
       `r (s * A[3, 1]) + (k * A[3, 1])` & `r (s * A[3, 2]) + (k * A[3, 2])` & `r (s * A[3, 3]) + (k * A[3, 3])` \\
       `r (s * A[4, 1]) + (k * A[4, 1])` & `r (s * A[4, 2]) + (k * A[4, 2])` & `r (s * A[4, 3]) + (k * A[4, 3])`
      \end{bmatrix}
    \end{split}
    \end{equation}

    ```{r}
s * A + k * A
    ```

1. $\mathbf{A} - \mathbf{B} = \mathbf{A} + \left(- \mathbf{B} \right)$

    ```{r}
A - B
A + (-B)
    ```

1. $\mathbf{A} - \mathbf{B} \neq \mathbf{B} - \mathbf{A}$

    ```{r}
A - B
B - A
    ```


Multiplication

```{r, echo = FALSE}
a <- c(
  4,
  1,
  -4,
  5,
  0,
  2
)
b <- c(
  8,
  3,
  -2,
  -5,
  3,
  1
)
A <- matrix(
  data = a,
  ncol = 3
)
B <- matrix(
  data = b,
  ncol = 2
)
```


```{r}
A %*% B
```
